---
title: "abc transporters pathway"
output: html_document
date: "2024-07-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(RColorBrewer)
library(ggh4x)
library(stringr)
```

```{r}
rootdir = "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/"
```

Figure 4: metabolic potential of autotrophic MAGs
```{r message=FALSE, warning=FALSE}
DRAM_KOs = read_csv(paste(rootdir, "DRAM_tables/DRAM_completeauto_OM_KO.csv", sep = ""))
DRAM_pathways = read_csv(paste(rootdir, "DRAM_tables/DRAM_completeauto_OM_pathway.csv", sep = ""), col_names = TRUE)
APEM_info = read_csv(paste(rootdir, "refined_analysis/refined_gtdb/gtdb_summary_06152026.csv", sep = ""))
auto = read_csv(paste(rootdir, "refined_analysis/refined_auto_MAGs_edited_06152026.csv", sep = ""))
catabolism_MAGs = read_csv(paste(rootdir, "refined_analysis/refined_tables/catabolism_MAGs_06152026.csv", sep = ""), col_names = "genome")
```

read in all the annotations, merge them with DRAM information and save as new df
```{r message=FALSE, warning=FALSE}
filenames = list.files(path = "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/refined_kofamscan/csv_processed/")
setwd("/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/refined_kofamscan/csv_processed/")

for (f in filenames){
  genome<-read_csv(f, col_names = TRUE)
  genome = as.data.table(genome$KO)
  genome$present = 1
  genome = genome %>%
    rename(!!f := "present")
  colnames(genome)[2] <-  sub("\\_processed.csv$", "", colnames(genome)[2])
  genome = genome %>%
    rename("gene_id" = "V1")
  genome = subset(genome, !duplicated(subset(genome, select=gene_id)))
  genome = left_join(DRAM_KOs, genome, by = "gene_id")
  assign(paste("genome", f, sep="_"), genome)
}

```

merge each genomes df's together into one giant matrix of gene presence/absence and clean up
```{r}
#get all data frames in list
df_list = mget(ls(pattern = "genome_O"))
#clean up the names so its just the bin ID
names(df_list) <- sub("\\_processed.csv$", "", names(df_list))
#merge all the dataframes together
a = do.call("cbind", df_list)

colnames(a)
#clean up df by renaming some files and dropping duplicate columns
a = a %>%
    rename("KO" = "genome_OC2_150m_MAG_100.fa.gene_id")
ab <- a %>% select(-contains(".gene_id"))

#reorder the KO columns so theyre in the same order for when you do cbind since that doesnt look for matches. had to do it this way since merge wasnt working (gave too many rows)
DRAM_pathways = DRAM_pathways[order(DRAM_pathways$gene_id,decreasing=TRUE),]
ab = ab[order(ab$KO,decreasing=TRUE),]
ab = cbind(ab, DRAM_pathways)

#dropping duplicate rows, renaming column headers, and replacing NAs with zeros
ab <- ab %>% select(-contains("gene_id"))
ab[is.na(ab)] <- 0
colnames(ab) = gsub("[.]R\\d+m\\d+m*[.]\\d+", "", colnames(ab))
```

manipulate presence/absence dataframe above to long form so I can add in the taxonomic identity and subset based on checkM quality metrics
```{r}
#now turn ab dataframe into long form so we can add tax and plot as heatmap
ab_long = ab %>% pivot_longer(cols = starts_with("genome_"), names_to = "genome", values_to = "present")
#remove the genome_ part so that it matches the data in the drep_tax df that I need to merge it with
ab_long$genome = sub("genome_", "", ab_long$genome)
ab_long$genome <- sub("\\.fa.*", "", ab_long$genome)

#modify gtdb bin names to match new refined bin names
ab_tax = left_join(ab_long, APEM_info, by = "genome")

ab_tax <- ab_tax %>%
  mutate(genome = str_replace(genome, "^(.*?)\\.fa\\.\\1\\.fa$", "\\1"))

#now use that list to pull out the mags of interest from the whole dataset
auto_MAGs = subset(ab_tax, (genome %in% catabolism_MAGs$genome))

length(unique(auto_MAGs$genome))
length(unique(auto$genome))
```

Figure 4 - cleaning up df for plotting
```{r}
auto_MAGs_simple = filter(auto_MAGs, auto_MAGs$header != "C1")
auto_MAGs_simple = filter(auto_MAGs, auto_MAGs$header != "C1-methane")
auto_MAGs_simple = filter(auto_MAGs_simple, auto_MAGs_simple$header != "Photosynthesis")
auto_MAGs_simple = filter(auto_MAGs_simple, auto_MAGs_simple$header != "Oxygen")

auto_MAGs_simple$phylum = sub("\\;c.*", "", auto_MAGs_simple$classification)
auto_MAGs_simple$class = sub("\\;o.*", "", auto_MAGs_simple$classification)
auto_MAGs_simple$kingdom = sub("\\;p.*", "", auto_MAGs_simple$classification)
auto_MAGs_simple$phylum = sub(".*p__", "", auto_MAGs_simple$phylum)
auto_MAGs_simple$class = sub(".*c__", "", auto_MAGs_simple$class)
auto_MAGs_simple$kingdom = sub(".*d__", "", auto_MAGs_simple$kingdom)
auto_MAGs_simple = auto_MAGs_simple %>% mutate(kingdom = ifelse(kingdom == "Bacteria", "B", kingdom))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(kingdom = ifelse(kingdom == "Archaea", "A", kingdom))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(phylum = ifelse(class == "Gammaproteobacteria", "Gammaproteobacteria", phylum))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(phylum = ifelse(class == "Alphaproteobacteria", "Alphaproteobacteria", phylum))

auto_MAGs_simple$label = paste(auto_MAGs_simple$kingdom, auto_MAGs_simple$phylum, sep = "_")

auto_MAGs_simple$header = replace(auto_MAGs_simple$header, auto_MAGs_simple$header=="carbon fixation", "Carbon Fixation")

auto_MAGs_simple = filter(auto_MAGs_simple, auto_MAGs_simple$present ==1)

auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "3-Hydroxypropionate bi-cycle", "A_3HP bi-cycle", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "rTCA Cycle", "A_rTCA", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "Calvin Cycle", "A_CBB", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "Hydroxypropionate-hydroxybutylate cycle", "A_3HP/4HB", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "Dicarboxylate-hydroxybutylate cycle", "A_DC/4HB", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(module == "Wood-Ljungdahl pathway", "A_WL", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(header == "methanotrophy", "CH4", header))
auto_MAGs_simple = auto_MAGs_simple %>% mutate(header = ifelse(header == "CO oxidation", "CO", header))

#drop K00441 since its not energy producing
auto_MAGs_simple = auto_MAGs_simple[-(which(auto_MAGs_simple$KO == "K00441")),]
```

```{r}
#percentage of MAGs with ABC tranporters associated with organic carbon

#Oligosaccaride
#Monosaccharide
Oligosaccaride = filter(auto_MAGs_simple, auto_MAGs_simple$module == "Oligosaccaride, polyol, lipid transporter")
Oligosaccaride_MAGs = aggregate(present ~ genome, data=Oligosaccaride, FUN=sum)
#29 genomes

Monosaccharide = filter(auto_MAGs_simple, auto_MAGs_simple$module == "Monosaccharide transporter")
Monosaccharide_MAGs = aggregate(present ~ genome, data=Monosaccharide, FUN=sum)
#35 genomes

shared_genomes <- intersect(Monosaccharide_MAGs$genome, Oligosaccaride_MAGs$genome)
#24 have both

merged_df <- merge(Monosaccharide_MAGs, Oligosaccaride_MAGs, by = "genome", all = TRUE)

# Step 2: Add columns indicating presence in each dataframe
merged_df$mono <- !is.na(merged_df$present.x) # TRUE if present in df1
merged_df$oligo <- !is.na(merged_df$present.y) # TRUE if present in df2

# Step 3: Collapse to unique genomes and retain source information
collapsed_df <- merged_df %>%
  select(genome, mono, oligo)

#40 have at least one
```

repeat for AA's
```{r}
#percentage of MAGs with ABC tranporters associated with organic carbon
BCAA = filter(auto_MAGs_simple, auto_MAGs_simple$module == "Branched-chain amino acid degradation")
BCAA_MAGs = aggregate(present ~ genome, data=BCAA, FUN=sum)
#122 genomes

AA = filter(auto_MAGs_simple, auto_MAGs_simple$module == "Phosphate and amino acid transporter")
AA_MAGs = aggregate(present ~ genome, data=AA, FUN=sum)
#98 genomes

peptide = filter(auto_MAGs_simple, auto_MAGs_simple$module == "Peptide and nickel transporter")
peptide_MAGs = aggregate(present ~ genome, data=peptide, FUN=sum)

shared_genomes <- Reduce(intersect, list(peptide_MAGs$genome, AA_MAGs$genome, BCAA_MAGs$genome))
#29 have all three

merged_df <- merge(AA_MAGs, BCAA_MAGs, by = "genome", all = TRUE)
merged_df = merge(merged_df, peptide_MAGs, by = "genome", all = TRUE)

# Step 2: Add columns indicating presence in each dataframe
merged_df$AA <- !is.na(merged_df$present.x) # TRUE if present in df1
merged_df$BCAA <- !is.na(merged_df$present.y) # TRUE if present in df2
merged_df$peptide <- !is.na(merged_df$present) # TRUE if present in df2

# Step 3: Collapse to unique genomes and retain source information
collapsed_df <- merged_df %>%
  select(genome, AA, BCAA, peptide)

#125 have at least one
```

```{r}
AA = filter(auto_MAGs_simple, auto_MAGs_simple$header == "abc transporter")
AA_MAGs = aggregate(present ~ genome, data=AA, FUN=sum)

genomes = unique(auto_MAGs_simple$genome)
ABC = unique(AA_MAGs$genome)

x = setdiff(genomes, ABC)
print(x)
#OC3_500m_MAG_167
##d__Bacteria;p__Actinobacteriota;c__Acidimicrobiia;o__Acidimicrobiales;f__MedAcidi-G1;g__UBA9410;s__UBA9410 sp014237715
```

```{r}
unique(auto_MAGs_simple$module)
glycine = filter(auto_MAGs_simple, auto_MAGs_simple$module == "glycine cleavage system")
glycine_MAGs = aggregate(present ~ genome, data=glycine, FUN=sum)
#120 MAGs
```

```{r}
phylum_plot <- ggplot(
  auto_MAGs_simple,
  aes(
    x = KO,
    y = genome,
    fill = module,
    alpha = present,
    width = 1,
    height = 1
  )
) +
  geom_tile(colour = NA, linewidth = 0.1) +

  facet_grid(
    rows = vars(label),
    cols = vars(header),
    scales = "free",
    space = "free",
    labeller = label_wrap_gen(width = 10)
  ) +

  scale_alpha_identity(guide = "none") +

  labs(
    x = "KO Number",
    y = "Putative Autotrophic MAGs"
  ) +

  scale_y_discrete(drop = TRUE) +

  theme_bw(base_size = 20) +
  theme(
    panel.grid = element_blank(),

    # X axis
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 24
    ),

    axis.text.y = element_blank(),

    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),

    axis.title.x = element_text(
      size = 30,
      face = "bold",
      margin = margin(t = 15)
    ),

    axis.title.y = element_text(
      size = 30,
      face = "bold",
      margin = margin(r = 15)
    ),

    strip.background = element_rect(
  fill = "white",
  colour = "black",
  linewidth = 0.4
),

    strip.text = element_text(
      size = 22,
      face = "bold",
      margin = margin(6, 6, 6, 6)
    ),
    
    strip.text.y.left = element_text(
      angle = 0,
      size = 22,
      face = "bold",
      hjust = 0,
      vjust = 0.5,
      margin = margin(r = 12)
    ),
    
    strip.text.y.right = element_text(
      angle = 0,
      size = 22,
      face = "bold",
      hjust = 0,
      vjust = 0.5,
      margin = margin(l = 12)
    ),
    
    strip.placement = "outside",
    strip.clip = "off",

    panel.spacing = unit(1, "lines"),

    panel.border = element_rect(
      colour = "black",
      linewidth = 0.4
    ),

    legend.position = "bottom",

    legend.title = element_text(
      size = 22,
      face = "bold"
    ),

    legend.text = element_text(
      size = 20
    ),

    plot.margin = margin(10, 20, 10, 20)
  )

ggsave(
  paste0(
    rootdir,
    "refined_analysis/refined_figures/06152026/refined_OMtransporters_06152026.pdf"
  ),
  plot = phylum_plot,
  width = 50,
  height = 50,
  units = "in",
  limitsize = FALSE
)
```

