---
title: "OC17_MAGs_refined_annotation"
output: html_document
date: "2024-06-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Libraries
```{r message=FALSE, warning=FALSE}
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(RColorBrewer)
library(ggh4x)
```

#Define Directories, Rename Bins
```{r}
rootdir = "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/"
oldgenomedir = paste(rootdir, "refined_analysis/refined_bins/", sep = "")
seqdir = "/scratch/users/rsalcedo/OC17wc_deepauto/sequences/"
```

```{R}
bin_list = gsub("_contigs.fa", "", list.files(path = oldgenomedir))
head(bin_list)
#rename to match

site_map <- c(
  "500"  = "OC2",
  "3000" = "OC3",
  "3500" = "OC4",
  "4000" = "OC5",
  "4500" = "OC6"
)

# Extract pieces and rename
renamed_binlist <- sapply(bin_list, function(x) {
  
  # Extract site, depth, and MAG number
  parts <- sub(
    "R([0-9]+)m([0-9]+)m\\.([0-9]+)",
    "\\1_\\2_\\3",
    x
  )
  
  pieces <- strsplit(parts, "_")[[1]]
  
  site  <- site_map[pieces[1]]
  depth <- pieces[2]
  mag   <- pieces[3]
  
  paste0(site, "_", depth, "m_MAG_", mag, ".fa")
})

renamed_binlist
```

```{r make new directory using jaffe refined bins}
# Directories
source_dir <- "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/jaffe_refined_genomes"
target_dir <- "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/06152026_refined_genomes"

if (!dir.exists(target_dir)) {
  dir.create(target_dir, recursive = TRUE)
}

# Full paths
source_files <- file.path(source_dir, renamed_binlist)
target_files <- file.path(target_dir, renamed_binlist)

# Check which files exist
missing_files <- renamed_binlist[!file.exists(source_files)]

if (length(missing_files) > 0) {
  warning(
    paste(
      "These files were not found:",
      paste(missing_files, collapse = ", ")
    )
  )
}

# Copy existing files
file.copy(
  from = source_files[file.exists(source_files)],
  to = target_files[file.exists(source_files)],
  overwrite = TRUE
)

# Report result
cat(
  "Copied",
  sum(file.exists(source_files)),
  "of",
  length(renamed_binlist),
  "files into",
  target_dir,
  "\n"
)
```

```{r redefine genome directories & save binlist}
genomedir <- "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/06152026_refined_genomes/"
renamed_binlist = list.files(genomedir)
#writeLines(renamed_binlist, paste0(rootdir, "/refined_binlist.txt"))
renamed_binlist = as.character(readLines(paste0(rootdir, "/refined_binlist.txt")))
```

#CheckM
```{r checkM lineage}
#make sure to activate base_metagenomics conda environment
sink(paste(rootdir, "sh_scripts/refined_checkM.sh", sep = ""))

cat("sbatch -J RcheckM -c 32 --mem-per-cpu 16G -p serc -t 160:00:00 --wrap 'checkm lineage_wf -t 32 --pplacer_threads 32 -x fa --tab_table ", genomedir, " ", rootdir, "refined_checkM/'", "\n", sep = "")

sink()

file.show(paste(rootdir, "sh_scripts/refined_checkM.sh", sep = ""))
```

```{r checkM qa}
sink(paste(rootdir, "sh_scripts/refined_checkM_qa.sh", sep = ""))
cat("sbatch -J Rqa -c 20 --mem-per-cpu 8G -p serc -t 160:00:00 --wrap 'checkm qa -t 20 -o 2 -f ", rootdir, "refined_checkM/refined_output_table.txt --tab_table ", rootdir, "refined_checkM/lineage.ms ", rootdir, "refined_checkM/'", "\n", sep = "")

sink()

file.show(paste(rootdir, "sh_scripts/refined_checkM_qa.sh", sep = ""))
```

```{r parsing checkM}
checkm = read_tsv("/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/refined_checkM/refined_output_table.txt")

checkm_mini = checkm %>% select(c("Bin Id", "Completeness", "Contamination"))

colnames(checkm_mini) = c("genome", "completeness", "contamination")

#write.csv(checkm_mini, paste0(rootdir, "refined_analysis/refined_checkM/refined_checkM_06152026.csv"), row.names = FALSE)
```

#Annotate Genomes
```{r prodigal command}
sink(paste(rootdir, "/sh_scripts/refined_prodigal.sh", sep = ""))
for (i in renamed_binlist) {
  cat("sbatch -J ", i, " -c 20 --mem-per-cpu 8G -p serc -t 160:00:00 --wrap 'prodigal -i ", genomedir, i, " -d ", rootdir, "refined_analysis/refined_prodigal/fna/", i, "_prodigalORFs.fna -a ", rootdir, "refined_analysis/refined_prodigal/faa/", i, "_prodigalORFs.faa -o ", rootdir, "refined_analysis/refined_prodigal/log/", i, "_prodigal.log'\n", sep = "")
  }
sink()

file.show(paste(rootdir, "sh_scripts/refined_prodigal.sh", sep = ""))
```

```{r kofamscan command}
#remember to activate kofamscan conda environment

kofamscan_path = "/home/groups/dekas/rebecca/kofamscan/kofam_scan-1.3.0/exec_annotation"
ko_list = "/home/groups/dekas/rebecca/kofamscan/ko_list"
ko_profiles = "/home/groups/dekas/rebecca/kofamscan/profiles"

sink(paste(rootdir, "sh_scripts/refined_kofamscan.sh", sep = ""))
cat("#!/bin/bash\n", "#SBATCH --job-name=koscan\n", "#SBATCH -p serc\n", "#SBATCH --time=168:00:00\n", "#SBATCH --ntasks=1\n", "#SBATCH --cpus-per-task=32\n", "#SBATCH --mem-per-cpu=8G\n\n", "for i in $(cat ", paste(rootdir, "refined_binlist.txt", sep = ""), ")\n", "do\n", kofamscan_path, " -o ", rootdir, "refined_analysis/refined_kofamscan/txt/${i}_KEGGIDs.txt -k ",  ko_list, " -p ", ko_profiles, " --cpu 32 --tmp-dir ", rootdir, "tmpdirs/${i}_tmpdir/ ", rootdir, "refined_analysis/refined_prodigal/faa/${i}_prodigalORFs.faa\n", "done\n", sep = "")
sink()

file.show(paste(rootdir, "sh_scripts/refined_kofamscan.sh", sep = ""))
```

```{r kofamparse command}
#remember to activate base_metagenomics conda environment and do ml ruby

kofamparse_path = "/home/groups/dekas/software/kofamparse/kofamparse"
sink(paste(rootdir, "sh_scripts/refined_kofamparse.sh", sep = ""))
  cat("#!/bin/bash\n", "#SBATCH --job-name=koparse\n", "#SBATCH -p serc\n", "#SBATCH --time=168:00:00\n", "#SBATCH --ntasks=1\n", "#SBATCH --cpus-per-task=32\n", "#SBATCH --mem-per-cpu=8G\n\n", "for i in $(cat ", paste(rootdir, "refined_binlist.txt", sep = ""), ")\n", "do\n", kofamparse_path, " ", rootdir, "refined_analysis/refined_kofamscan/txt/${i}_KEGGIDs.txt ", rootdir, "refined_analysis/refined_kofamscan/csv/${i}_KEGGIDs.csv\n", "done\n", sep = "")
sink()

file.show(paste(rootdir, "sh_scripts/refined_kofamparse.sh", sep = ""))

```

```{r kofamscan thresholding}
KO_thresholds = read_csv(paste(rootdir, "bitscore_thresholds.csv", sep = ""))

#define list of KO's im interested in
KOs = list("K00194", "K00197", "K00198", "K00855", "K01601", "K01602", "K08691", "K09709", "K14466", "K14467", "K14468", "K14469", "K14470", "K15016", "K15017", "K15023", "K15039", "K15052", "K15230", "K15231", "K15232", "K15233", "K15234", "K18861", "K14138")

for(i in renamed_binlist) {                                           
  kegg = read_csv(paste(rootdir, "refined_analysis/refined_kofamscan/csv/", i, "_KEGGIDs.csv", sep = ""), col_names = TRUE)
  kegg$score = as.integer(kegg$score)
  kegg$threshold = as.integer(kegg$threshold)
  kegg = rename(kegg, gene = "Gene name") 
  kegg = rename(kegg, KO = "KO number") 
  kegg = rename(kegg, evalue = "e-value") 
  kegg_besthit = kegg %>% 
    group_by(gene) %>% 
    slice_max(score)
  kegg_besthit = subset(kegg_besthit, kegg_besthit$evalue < 1e-6)
  kegg_besthit = subset(kegg_besthit, kegg_besthit$score >= threshold)
  
  auto_kegg = subset(kegg_besthit, (KO %in% KOs))
  for (x in KOs) {
    threshold = with(KO_thresholds, new[KO == x])
    assign(paste("bitscore_", x, sep = ""), threshold)
    pass <- auto_kegg[which(auto_kegg$KO == x & auto_kegg$score >= threshold), ]
    assign(paste("bitscore_pass_", x, sep = ""), pass)
  }
  auto_filtered <- rbind(bitscore_pass_K00194, bitscore_pass_K00197, bitscore_pass_K00198, bitscore_pass_K00855, bitscore_pass_K01601, bitscore_pass_K01602, bitscore_pass_K08691, bitscore_pass_K09709, bitscore_pass_K14466, bitscore_pass_K14467, bitscore_pass_K14468, bitscore_pass_K14469, bitscore_pass_K14470, bitscore_pass_K15016, bitscore_pass_K15017, bitscore_pass_K15023, bitscore_pass_K15039, bitscore_pass_K15052, bitscore_pass_K15230, bitscore_pass_K15231, bitscore_pass_K15232, bitscore_pass_K15233, bitscore_pass_K15234, bitscore_pass_K18861, bitscore_pass_K14138)
  auto_filtered  = auto_filtered  %>%
    mutate(pathway = case_when(
      endsWith(KO, "00855") ~ "CBB",
      endsWith(KO, "01601") ~ "CBB",
      endsWith(KO, "01602") ~ "CBB",
      endsWith(KO, "15230") ~ "rTCA", 
      endsWith(KO, "15231") ~ "rTCA",
      endsWith(KO, "15232") ~ "rTCA",
      endsWith(KO, "15233") ~ "rTCA",
      endsWith(KO, "15234") ~ "rTCA",
      endsWith(KO, "00198") ~ "WL",
      endsWith(KO, "14138") ~ "WL",
      endsWith(KO, "00197") ~ "WL",
      endsWith(KO, "00194") ~ "WL",
      endsWith(KO, "15023") ~ "WL",
      endsWith(KO, "14468") ~ "3HP Bicycle",
      endsWith(KO, "15052") ~ "3HP Bicycle",
      endsWith(KO, "08691") ~ "3HP Bicycle",
      endsWith(KO, "14469") ~ "3HP Bicycle",
      endsWith(KO, "14470") ~ "3HP Bicycle",
      endsWith(KO, "09709") ~ "3HP Bicycle",
      endsWith(KO, "15039") ~ "3HP/4HB",
      endsWith(KO, "14466") ~ "3HP/4HB",
      endsWith(KO, "15017") ~ "3HP/4HB or DC/4HB",
      endsWith(KO, "15016") ~ "3HP/4HB or DC/4HB",
      endsWith(KO, "18861") ~ "3HP/4HB or DC/4HB",
      endsWith(KO, "14467") ~ "DC/4HB"
  ))
  duprows <- rownames(kegg_besthit) %in% rownames(auto_filtered)
  processed = rbind(auto_filtered, kegg_besthit[!duprows,])
  write.csv(processed, paste(rootdir, "refined_analysis/refined_kofamscan/csv_processed/", i, "_processed.csv", sep = ""))
}

```

#gtdbtk
```{r generate command}
#download db 
#wget https://data.gtdb.ecogenomic.org/releases/release207/207.0/auxillary_files/	gtdbtk_r207_v2_data.tar.gz
#using release version v207

#switch base to base_metagenomics conda env
sink(paste(rootdir, "/sh_scripts/gtdb-tk.sh", sep = ""))

cat("sbatch -J gtdbtk -c 32 --mem-per-cpu 8G -p serc -t 160:00:00 --wrap 'gtdbtk classify_wf --genome_dir ", genomedir, " --out_dir ", rootdir, "refined_analysis/refined_gtdb/ -x fa --cpus 32 --skip_ani_screen'", "\n", sep = "")
sink()

file.show(paste(rootdir, "sh_scripts/gtdb-tk.sh", sep = ""))

```

```{r parse gtdb results}
gtdb_dir = paste0(rootdir, "refined_analysis/refined_gtdb")

archaea=read_tsv(file.path(gtdb_dir, "gtdbtk.ar53.summary.tsv"))
bacteria=read_tsv(file.path(gtdb_dir, "gtdbtk.bac120.summary.tsv"))

gtdb_combined=bind_rows(archaea, bacteria)
gtdb_summary = gtdb_combined %>% select(c("user_genome", "classification"))
colnames(gtdb_summary) = c("genome", "classification")
write.csv(gtdb_summary, paste0(rootdir, "/refined_analysis/refined_gtdb/gtdb_summary_06152026.csv"), row.names = FALSE)
```

# Assess metabolic potential of autotrophic MAGs (make table for Fig. 4)
```{r read in all tables needed}
DRAM_pathways = read_tsv(paste(rootdir, "DRAM_tables/DRAM_energy_pathways_v5.tsv", sep = ""), col_names = TRUE)
DRAM_KOs = DRAM_pathways %>% select(gene_id)
checkM_info = checkm_mini
gtdb_info = gtdb_summary
```

read in all the annotations, merge them with DRAM information and save as new df
```{r message=FALSE, warning=FALSE}
filenames = list.files(path = "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/refined_kofamscan/csv_processed/")
setwd("/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/refined_analysis/refined_kofamscan/csv_processed/")

for (f in filenames) {

  sample_name <- sub("_processed\\.csv$", "", basename(f))

  genome <- read_csv(f, col_names = TRUE, show_col_types = FALSE) %>%
    select(KO) %>%
    distinct(KO) %>%
    transmute(
      gene_id = KO,
      present = 1
    )

  names(genome)[names(genome) == "present"] <- sample_name

  genome <- left_join(DRAM_KOs, genome, by = "gene_id")

  assign(
    paste0("genome_", sample_name),
    genome
  )
}
```

merge each genomes df's together into one giant matrix of gene presence/absence and clean up
```{r}
#get all data frames in list
df_list = mget(ls(pattern = "genome_O"))
#clean up the names so its just the bin ID
names(df_list) <- sub("\\_processed.csv$", "", names(df_list))
#merge all the dataframes together
a = do.call("cbind", df_list)

#clean up df by renaming some files and dropping duplicate columns
a = a %>%
    dplyr::rename("KO" = "genome_OC2_150m_MAG_100.fa.gene_id")
ab <- a %>% select(-contains(".gene_id"))

#reorder the KO columns so theyre in the same order for when you do cbind since that doesnt look for matches. had to do it this way since merge wasnt working (gave too many rows)
DRAM_pathways = DRAM_pathways[order(DRAM_pathways$gene_id,decreasing=TRUE),]
ab = ab[order(ab$KO,decreasing=TRUE),]
ab = cbind(ab, DRAM_pathways)

#dropping duplicate rows, renaming column headers, and replacing NAs with zeros
ab <- ab %>% select(-contains("gene_id"))
ab[is.na(ab)] <- 0
colnames(ab) = gsub("[.]R\\d+m\\d+m*[.]\\d+", "", colnames(ab))

colnames(ab)
```

manipulate presence/absence dataframe above to long form so I can add in the taxonomic identity and subset based on checkM quality metrics
```{r}
#now turn ab dataframe into long form so we can add tax and plot as heatmap
ab_long = ab %>% pivot_longer(cols = starts_with("genome_"), names_to = "genome", values_to = "present")
#remove the genome_ part so that it matches the data in the drep_tax df that I need to merge it with
ab_long$genome = sub("genome_", "", ab_long$genome)
ab_long$genome = sub("^(.*)\\.fa\\.\\1\\.fa$", "\\1", ab_long$genome)

#merge the dataframes
ab_tax = left_join(ab_long, gtdb_info, by = "genome")

#clean up taxonomy names for plotting
ab_tax$phylum = sub("\\;c.*", "", ab_tax$classification)
ab_tax$class = sub("\\;o.*", "", ab_tax$classification)

#add in checkM info
ab_tax = left_join(ab_tax, checkM_info, by = "genome")
#drop MAGs with over 5% contamination and under 50% completeness
ab_tax_505 = filter(ab_tax, ab_tax$completeness >= 50 & ab_tax$contamination <= 5)

#make a list of genomes with one or more
#start by getting only rows of autotrophy genes
auto = filter(ab_tax_505, ab_tax_505$header == "carbon fixation")
#sum that data by genome to find the total carbon fixation genes in each genome
aa = aggregate(present ~ genome, data=auto, FUN=sum)
#subset the above df to get a list of genomes that have at least two genes involved in carbon fixation
auto_sum = filter(aa, aa$present >= 1)
#now use that list to pull out the mags of interest from the whole dataset
auto_MAGs = subset(ab_tax_505, (genome %in% auto_sum$genome))

length(unique(auto_MAGs$genome))
length(unique(ab_tax_505$genome))

write.csv(auto_MAGs, paste(rootdir, "refined_analysis/refined_auto_MAGs_06152026.csv", sep = ""), row.names = FALSE)
```

re-do gene trees
update that file with results of gene trees (OC17_refinedMAG_gene_tree_06152026.Rmd)
```{r}
auto_MAGs = read.csv(paste(rootdir, "refined_analysis/refined_auto_MAGs_06152026.csv", sep = ""))

auto_MAGs$gene_description <- gsub("\\[EC([0-9])", "[EC:\\1", auto_MAGs$gene_description)

auto_MAGs_distinct <- distinct(auto_MAGs)

colnames(auto_MAGs_distinct)
```

```{r}
remove = read.csv(paste0(rootdir, "refined_analysis/gene_trees/06152026_genetree_to_remove.csv"))
remove$X = NULL
colnames(remove)
```

```{r}
#Add the target 'present' value to the remove dataframe
remove_mod <- remove %>% 
  mutate(present = 0)

auto_MAGs_edited <- auto_MAGs_distinct %>%
  # Left join and create a flag column from the 'remove' df
  left_join(
    remove %>% mutate(to_remove = TRUE), 
    by = c("genome", "KO", "module", "header")
  ) %>%
  # If it matched, change present to 0; otherwise keep the original value
  mutate(present = if_else(!is.na(to_remove), 0, present)) %>%
  # Drop the temporary flag column
  select(-to_remove)

nrow(anti_join(auto_MAGs_distinct, auto_MAGs_edited))
```

```{r}
write.csv(auto_MAGs_edited, paste0(rootdir, "/refined_analysis/refined_auto_MAGs_edited_06152026.csv"))
```

```{r}
auto_MAGs_edited = read.csv(paste0(rootdir, "/refined_analysis/refined_auto_MAGs_edited_06152026.csv"))
```

Table S6: genome, taxonomy, size, GC, carbon fixation pathway present, completeness, contamination, sample generated from, relative abundance, site, depth
```{r}
tableS6 = auto_MAGs_edited
tableS6 = filter(tableS6, tableS6$present == 1)

CBB <- c("K01602", "K00855", "K01601")
HP <- c("K14468", "K15052", "K08691", "K14469", "K14470", "K14449", "K09709")
rTCA = c("K15230", "K15231", "K15232", "K15233", "K15234")
WL = c("K00198", "K14138", "K00197", "K00194", "K15023")
HP_HB = c("K15039", "K14466")
HP_DC_HB = c("K15017", "K15016", "K18861", "K14467")

tableS6$pathway = "X"
tableS6 <- tableS6 %>%
  group_by(genome) %>%
  mutate(pathway = case_when(
    any(KO %in% CBB) & any(KO %in% HP) ~ "3HP bicycle, CBB",
    any(KO %in% CBB) ~ "CBB",
    any(KO %in% HP) ~ "3HP bi-cycle",
    any(KO %in% rTCA) ~ "rTCA",
    any(KO %in% WL) ~ "WL",
    any(KO %in% HP_HB) ~ "3HP/4HB",
    any(KO %in% HP_DC_HB) ~ "3HP/4HB or DC/4HB",
    TRUE ~ NA_character_
  ))

tableS6$KO = NULL
tableS6$gene_description = NULL
tableS6$header = NULL
tableS6$module = NULL
tableS6$present = NULL

tableS6 = subset(tableS6,!is.na(pathway))

tableS6 = distinct(tableS6, genome, .keep_all = TRUE)

#modify pathway based on taxonomy
tableS6 <- tableS6 %>%
  mutate(
    pathway = case_when(
      grepl("halobacteriota", phylum, ignore.case = TRUE) & pathway == "3HP/4HB or DC/4HB" ~ "DC/4HB",
      grepl("thermoproteota|thermoplasmatota|planctomycetota|myxococcota", phylum, ignore.case = TRUE) & pathway == "3HP/4HB or DC/4HB" ~ "3HP/4HB",
      TRUE ~ pathway
    )
  )

tableS6 <- tableS6 %>%
  separate(
    col = genome,         
    into = c("site_generated", "depth_generated"),
    sep = "_",                    
    remove = FALSE                
  )

#salcedo_genomes = tableS6$genome
#write.csv(salcedo_genomes, paste0(rootdir, "/refined_analysis/refined_tables/salcedo_genomes_06152026.csv"))

#specify if genome was published in alex's or nestor's papers
#read in table from alex for previously reported 
prev_reported = read.csv(paste0(rootdir, "refined_analysis/jaffe2026_refined_genome_metadata.csv"))
nestor_genomes = prev_reported$genome_name[prev_reported$reported_in == "Arandia-Gorostidi et al. 2024"]
jaffe_genomes = prev_reported$genome_name[prev_reported$reported_in == "Jaffe et al. 2025"]

# Add the previously_reported column
tableS6 <- tableS6 %>%
  mutate(previously_reported = case_when(
    genome %in% jaffe_genomes  ~ "Jaffe et al., 2025",
    genome %in% nestor_genomes ~ "Arandia-Gorostidi et al. 2024", 
    TRUE                       ~ "no"
  ))

#using info made in OC17_refinedMAGs_figure3_06152026.Rmd analysis load in the catabolism MAGs
catabolism_MAGs = read.csv(paste0(rootdir, "/refined_analysis/refined_tables/catabolism_MAGs_06152026.csv"))

catabolism_MAGs = as.character(catabolism_MAGs$x)

tableS6 <- tableS6 %>%
  mutate(catabolism_found = ifelse(genome %in% catabolism_MAGs, "TRUE", "FALSE"))

#add in genome length from coverM
coverm_raw = read.csv(paste0(rootdir, "refined_analysis/merged_coverM/refined_coverM_all_06152026.csv"))

coverm_mini = distinct(coverm_raw, genome, .keep_all = TRUE) %>% select(c("genome", "Length"))
tableS6 = merge(tableS6, coverm_mini, by = "genome")

tableS6$phylum = sub("\\;c.*", "", tableS6$classification)
tableS6$class = sub("\\;o.*", "", tableS6$classification)
tableS6$kingdom = sub("\\;p.*", "", tableS6$classification)
tableS6$phylum = sub(".*p__", "", tableS6$phylum)
tableS6$class = sub(".*c__", "", tableS6$class)
tableS6$kingdom = sub(".*d__", "", tableS6$kingdom)
tableS6 = tableS6 %>% mutate(kingdom = ifelse(kingdom == "Bacteria", "B", kingdom))
tableS6 = tableS6 %>% mutate(kingdom = ifelse(kingdom == "Archaea", "A", kingdom))
tableS6 = tableS6 %>% mutate(phylum = ifelse(class == "Gammaproteobacteria", "Gammaproteobacteria", phylum))
tableS6 = tableS6 %>% mutate(phylum = ifelse(class == "Alphaproteobacteria", "Alphaproteobacteria", phylum))

tableS6$label = paste(tableS6$kingdom, tableS6$phylum, sep = "_")

tableS6$kingdom = NULL
tableS6$class = NULL
tableS6$phylum = NULL
tableS6$X = NULL

write.csv(tableS6, paste(rootdir, "refined_analysis/refined_tables/refined_tableS6_06152026.csv", sep = ""), row.names = FALSE)
```

Table S10: genome, KO, KO description, larger metabolic function, site, depth
```{r}
#read in updated auto_MAGs after manual refinement of catabolic genes using trees (to become tableS10)
auto_MAGs = read_csv(paste(rootdir, "refined_analysis//refined_auto_MAGs_edited_06152026.csv", sep = ""))

tableS10 = auto_MAGs
tableS10 = filter(tableS10, tableS10$present >= 1)
tableS10 = select(tableS10, "genome", "KO", "gene_description", "module")
tableS10 = dplyr::rename(tableS10, KO_description = gene_description)
tableS10 = dplyr::rename(tableS10, metabolic_function = module)
tableS10$present = 1

write.csv(tableS10, paste(rootdir, "refined_analysis/refined_tables/refined_tableS10_06152026.csv", sep = ""), row.names = FALSE)
```

#determine relative abundance of genomes
map to whole drep set but replace old mags with refined versions before running coverM
```{r make new coverM genome set directory}
# Directories
orig_dir = paste0(rootdir, "/dereplicated_genomes/")
refined_dir = genomedir
out_dir = paste0(rootdir, "/coverM_genome_set")
```

```{r rename unrefined genomes to match naming scheme}
genome_dir <- "/scratch/users/rsalcedo/OC17wc_deepauto/genome_based/dereplicated_genomes"

# Site conversion map
site_map <- c(
  "500"  = "OC2",
  "3000" = "OC3",
  "3500" = "OC4",
  "4000" = "OC5",
  "4500" = "OC6"
)

# Get fasta files
files <- list.files(
  genome_dir,
  pattern = "\\.fa$",
  full.names = TRUE
)

# Convert names
new_names <- sapply(basename(files), function(x) {
  
  # remove .fa
  base <- sub("\\.fa$", "", x)
  
  parts <- sub(
    "[AR]([0-9]+)m([0-9]+)m\\.([0-9]+)",
    "\\1_\\2_\\3",
    base
  )
  
  pieces <- strsplit(parts, "_")[[1]]
  
  site  <- site_map[pieces[1]]
  depth <- pieces[2]
  mag   <- pieces[3]
  
  paste0(site, "_", depth, "m_MAG_", mag, ".fa")
})

# Preview changes
rename_table <- data.frame(
  old = basename(files),
  new = new_names
)

print(rename_table)

# actually rename files
file.rename(
  from = files,
  to = file.path(genome_dir, new_names)
)
```

```{r create new coverM directory that is drep set + refined genomes}
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

orig_files <- list.files(orig_dir, pattern = "\\.fa$", full.names = TRUE)
refined_files <- list.files(refined_dir, pattern = "\\.fa$", full.names = TRUE)

# generate a match key
get_key <- function(x) {
  x <- basename(x)
  x <- sub("^[AR]", "", x)
  x <- sub("_contigs", "", x)
  x <- sub("\\.fa$", "", x)
  return(x)
}

#  lookup table for refined files
refined_keys <- get_key(refined_files)
refined_lookup <- setNames(refined_files, refined_keys)

for (f in orig_files) {

  key <- get_key(f)

  if (key %in% names(refined_lookup)) {

    refined_file <- refined_lookup[[key]]
    out_name <- sub("_contigs", "", basename(refined_file))

    file.copy(
      refined_file,
      file.path(out_dir, out_name),
      overwrite = TRUE
    )

  } else {

    file.copy(
      f,
      file.path(out_dir, basename(f)),
      overwrite = TRUE
    )

  }
}

# summary
n_replaced <- sum(get_key(orig_files) %in% names(refined_lookup))

cat("Original files:", length(orig_files), "\n")
cat("Refined files:", length(refined_files), "\n")
cat("Files replaced by refined versions:", n_replaced, "\n")
cat("Files in output directory:", length(list.files(out_dir)), "\n")
```

```{r define sample list for trimmed reads}
sample_list = c("3000m1000m", "3000m150m", "3000m2000m", "3000m3000m", "3000m500m", "3000m50m", "3500m1000m", "3500m150m", "3500m2000m", "3500m3000m", "3500m500m", "3500m50m", "4000m1000m", "4000m150m", "4000m2000m", "4000m3000m", "4000m500m", "4000m50m", "4500m150m", "4500m2000m", "4500m3000m", "4500m4000m", "4500m500m", "4500m50m", "500m150m", "500m500m", "500m50m", "4500m1000m")
```

```{r coverm command}
sink(paste(rootdir,"sh_scripts/merged_coverM.sh", sep = ""))
for (i in sample_list) {
  cat("sbatch -J ", i, "_coverM -c 20 --mem-per-cpu 8G -p serc -t 160:00:00 --wrap 'coverm genome --coupled ", seqdir, "trimmed_reads/", i, "_R1_trimmed.fastq.gz ", seqdir, "trimmed_reads/", i, "_R2_trimmed.fastq.gz --genome-fasta-directory ",  rootdir, "coverM_genome_set/ -x fa -t 20 --min-read-percent-identity 95 --output-format sparse --min-covered-fraction 0 -m count mean covered_fraction length -o ", rootdir, "refined_analysis/merged_coverM/", i, "_coverM.tsv'\n", sep = "")
  }
sink()

file.show(paste(rootdir,"sh_scripts/merged_coverM.sh", sep = ""))
```

```{r parsing coverM and merging message=FALSE, warning=FALSE}
for(i in sample_list) {                                           
  coverM = read_tsv(paste(rootdir, "refined_analysis/merged_coverM/", i, "_coverM.tsv", sep = ""), col_names = TRUE)
  coverM$Sample = gsub("_R1_trimmed.fastq.gz", "", coverM$Sample)
  coverM = dplyr::rename(coverM, genome = Genome)
  coverM$genome = gsub("_contigs", "", coverM$genome)
  coverM = dplyr::rename(coverM, covered_fraction = "Covered Fraction")
  assign(paste("coverM_", i, sep = ""), coverM)
}


df_list = mget(ls(pattern = "coverM_"))
coverM = rbindlist(df_list, use.names = TRUE)
coverM <- coverM %>%
  separate(Sample, into = c("site", "depth"), sep = "m")

site_map <- c(
  "500"  = "OC2",
  "3000" = "OC3",
  "3500" = "OC4",
  "4000" = "OC5",
  "4500" = "OC6"
)

coverM <- coverM %>%
  mutate(site = recode(site, !!!site_map))

write_csv(coverM, paste(rootdir, "refined_analysis/merged_coverM/refined_coverM_all_06152026.csv", sep = ""))
```
