---
title: "OC17 gene abundance and geochemistry"
output: html_document
date: "2026-07-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Set Up
```{r load libraries, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(RColorBrewer)
library(broom)
```

```{r define root dir}
rootdir = "/scratch/users/rsalcedo/OC17wc_deepauto/gene_based/geochem/"
abundir = "/scratch/users/rsalcedo/OC17wc_deepauto/gene_based/R_output/"
```

```{r read in geochemistry data}
geochem_original = read.csv(paste(rootdir, "arandia-gorostidi_etal_isme_supptables_urea_tableS1.csv", sep = ""))
geochem = geochem_original
geochem$sample = paste(geochem$site, "m", geochem$depth, "m", sep = "")

print(colnames(geochem))

geochem = select(geochem, c("sample", "site", "depth", "Salinity", "Temperature_C", "Fluorescence", "Dissolved_oxygen_sat.", "Average_urea_concentration_nM", "Ammonium_concentration_nM", "Average_nitrate_concentration_uM"))
```

```{r read in abundance data, message=FALSE, warning=FALSE}
pathways = read.csv(paste(abundir, "transect_pathwaydf.csv", sep = ""))
pathways$X = NULL
KOs = read.csv(paste(abundir, "transect_KOdf.csv", sep = ""))
KOs$X = NULL

pathway_no_sc = subset(pathways, pathways$metabolism!="single copy")

totals = pathway_no_sc %>% group_by(site_m, depth_m) %>% summarize(sum(normalized_cov))
totals = totals %>% 
       dplyr::rename("auto_normcov" = "sum(normalized_cov)")
totals$sample = paste(totals$site_m, "m", totals$depth_m, "m", sep = "")

pathway_no_sc = pathway_no_sc %>%
  dplyr::rename("auto_normcov" = "normalized_cov")
pathway_no_sc$sample = paste(pathway_no_sc$site_m, "m", pathway_no_sc$depth_m, "m", sep = "")
```

```{r merge df's}
pathway_geochem = merge(geochem, pathway_no_sc, by = "sample")
colnames(pathway_geochem)
pathway_geochem = select(pathway_geochem, c("sample", "site", "depth", "Salinity", "Temperature_C", "Fluorescence", "Dissolved_oxygen_sat.", "Average_urea_concentration_nM", "Ammonium_concentration_nM", "Average_nitrate_concentration_uM", "pathway", "auto_normcov"))


totals_geochem = merge(geochem, totals, by = "sample")
```

#Pearson's R Correlation
```{r List of variables to test}
variables = c("sample", "site", "depth", "Salinity", "Temperature_C", "Fluorescence", "Dissolved_oxygen_sat.", "Average_urea_concentration_nM", "Ammonium_concentration_nM", "Average_nitrate_concentration_uM")
```

```{r define function to calculate}
calculate_correlation = function(df) {
  results = data.frame(Variable = character(), 
                        Correlation = numeric(), 
                        P.Value = numeric(), 
                        Adjusted.P.Value = numeric(), 
                        stringsAsFactors = FALSE)
  
  for (var in variables) {
    x = df[[var]]
    y = df$auto_normcov
    
    valid_indices = complete.cases(x, y) & is.finite(x) & is.finite(y)
    x = x[valid_indices]
    y = y[valid_indices]
    
    if (length(unique(x)) > 1 && length(unique(y)) > 1) {
      print(paste("Processing variable:", var))
      test = try(cor.test(x, y, method = "pearson"), silent = TRUE)
      
      if (class(test) != "try-error" && !is.null(test$estimate)) {
        results = rbind(results, 
                         data.frame(Variable = var, 
                                    Correlation = test$estimate, 
                                    P.Value = test$p.value,
                                    Adjusted.P.Value = p.adjust(test$p.value, method = "bonferroni", n = length(variables)),
                                    stringsAsFactors = FALSE))
      } else {
        print(paste("Skipping variable:", var, "- correlation not computed"))
      }
    } else {
      print(paste("Skipping variable:", var, "- insufficient unique values"))
    }
  }
  
  return(results)
}
```

```{r total abundance - all depths}
tot_correlation_results = totals_geochem %>%
  group_by(site) %>%
  group_modify(~ calculate_correlation(.x)) %>%
  ungroup()

print(tot_correlation_results)

#write.csv(tot_correlation_results, paste0(rootdir, "/totals_correlation_results.csv"), row.names = FALSE)
```

```{r total - excluding 50m all sites}
totals_geochem_filtered = totals_geochem %>% filter(depth != 50)

tot50_correlation_results = totals_geochem_filtered %>%
  group_by(site) %>%
  group_modify(~ calculate_correlation(.x)) %>%
  ungroup()

print(tot50_correlation_results)

#write.csv(tot50_correlation_results, paste0(rootdir, "/totals50_correlation_results.csv"), row.names = FALSE)
```

```{r pathways - all depths}
correlation_results_pathways = pathway_geochem %>%
  group_by(pathway, site) %>%
  group_modify(~ calculate_correlation(.x)) %>%
  ungroup()

print(correlation_results_pathways)

#write.csv(correlation_results_pathways, paste0(rootdir, "/pathway_correlation_results.csv"), row.names = FALSE)
```

```{r pathways - excluding 50m}
pathway_geochem_filtered = pathway_geochem %>% filter(depth != 50)

correlation_results_pathways50 = pathway_geochem_filtered %>%
  group_by(pathway, site) %>%
  group_modify(~ calculate_correlation(.x)) %>%
  ungroup()

print(correlation_results_pathways50)

#write.csv(correlation_results_pathways50, paste0(rootdir, "/pathway50_correlation_results.csv"), row.names = FALSE)
```

#Welch's One Way ANOVA for meso v bathy
```{r define realms}
bathy = c("2000", "3000", "4000")
epi = c("50", "150")
meso = c("500", "1000")
```

```{r total}
totals_geochem_2 = totals_geochem %>% mutate(realm = case_when(
    depth %in% epi   ~ "epi",
    depth %in% meso  ~ "meso",
    depth %in% bathy ~ "bathy",
  ))

totals_geochem_2_filtered = totals_geochem_2 %>% filter(realm %in% c("bathy", "meso"))

anova_data = totals_geochem_2_filtered %>%
  filter(!is.na(realm)) %>%
  mutate(realm = as.factor(realm))

normcov_anova = oneway.test(auto_normcov ~ realm, data = anova_data, var.equal = FALSE)
summary(normcov_anova)
print(normcov_anova)

#which one is higher
anova_data %>%
  group_by(realm) %>%
  summarise(
    Mean_Normcov = mean(auto_normcov, na.rm = TRUE),
    Sample_Size = n()
  )
```

```{r pathways}
pathway_geochem_2 = pathway_geochem %>% mutate(realm = case_when(
    depth %in% epi   ~ "epi",
    depth %in% meso  ~ "meso",
    depth %in% bathy ~ "bathy",
  ))

pathway_geochem_2_filtered = pathway_geochem_2 %>% filter(realm %in% c("bathy", "meso"))

pathways = unique(pathway_geochem_2_filtered$pathway)
anova_data = pathway_geochem_2_filtered %>%
  mutate(realm = as.factor(realm))

pathway_results = data.frame()

for (pat in pathways) {
  pat_data = anova_data %>% filter(pathway == pat)
  if (length(unique(pat_data$realm)) == 2) {
    t_test = try(t.test(auto_normcov ~ realm, data = pat_data, var.equal = FALSE))
    if (!inherits(t_test, "try-error")) {
      means = pat_data %>%
        group_by(realm) %>%
        summarise(mean_value = mean(auto_normcov, na.rm = TRUE), .groups = "drop")
      
      realm_names = means$realm
      realm_means = means$mean_value
      
      pathway_results = rbind(pathway_results, data.frame(
        Pathway = pat,
        Group1 = realm_names[1],
        Group1_Mean = realm_means[1],
        Group2 = realm_names[2],
        Group2_Mean = realm_means[2],
        T_Statistic = t_test$statistic,
        P_Value = t_test$p.value,
        stringsAsFactors = FALSE
      ))
    }
  }
}

pathway_results = pathway_results %>%
  mutate(Adjusted_P_Value = p.adjust(P_Value, method = "bonferroni")) %>% 
  arrange(P_Value)

pathway_results
```

#Significant Results
Pearson's Correlation test with p value Bonferonni correction
  •	Correlations assessed within each site
    o 	Total abundance – all depths, all sites
      	OC6, fluorescence (0.97, 0.0012)
    o	Total abundance – no 50m, all sites
      	Nothing
    o	Pathway abundance – all depths, all sites
      	OC6: CBB, fluorescence (0.98, 0.0018)
    o	Pathway abundance – no 50m, all sites
      	OC6: rTCA, nitrate (-0.96, 0.015)
      	OC5: rTCA, nitrate (-0.98, 0.017)
Welch’s one way anova test (doesn’t assume equal variance or sample size across groups) *bathypelagic vs mesopelagic only – all ignore 50m and 150m
  •	Total abundance, all sites
    o	P value: 0.00476, higher in mesopelagic
  •	Pathway abundance, all sites
    o	CBB, p-value: 0.0013. higher in mesopelagic
